### 2020-10-29 bi-model.Rmd
asd

---
title: "ps4_models"
author: "James Bao"
date: "10/27/2020"
output: pdf_document
---

```{r}
library(haven)
library(tidyverse)
library(nnet)
library(Metrics)
library(brms)
```
```{r}
# remove.packages("rstan")

```

```{r}
# install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)

```

```{r}
raw_data <- read_dta("inputs/data/ns20200625/ns20200625.dta")
# Add the labels
raw_data <- labelled::to_factor(raw_data)
```
```{r}
colnames(raw_data)
```

```{r}
reduced_data <- 
  raw_data %>% 
  select(vote_intention,
         vote_2020,
         state,
         race_ethnicity,
         age,
         gender,
         household_income,
         education
  )

reduced_data <- reduced_data[complete.cases(reduced_data), ]

reduced_data <- reduced_data[reduced_data$vote_2020 == "Donald Trump" |
                             reduced_data$vote_2020 == "Joe Biden",]

reduced_data$race_ethnicity <- as.character(reduced_data$race_ethnicity)
reduced_data$race_ethnicity[
    reduced_data$race_ethnicity == 'Asian (Asian Indian)' |
    reduced_data$race_ethnicity == 'Asian (Korean)' |
    reduced_data$race_ethnicity == 'Asian (Filipino)' |
    reduced_data$race_ethnicity == 'Asian (Vietnamese)'|
    grepl("Pacific", reduced_data$race_ethnicity)
  ] <- ('Asian (Other)')
reduced_data$race_ethnicity <- as.factor(reduced_data$race_ethnicity)

reduced_data <- droplevels(reduced_data)

total_rows <- nrow(reduced_data)
table(reduced_data$vote_2020)

reduced_data
```
```{r}
shuffle_indices <- sample(total_rows)
data <- reduced_data[shuffle_indices,]
boundary <- as.integer(total_rows*0.9)

training <- data[0:boundary,]
model <- brm(formula = vote_2020 ~ (1|state) + (1|race_ethnicity) + (1|gender) + 
                          (1|household_income) + (1|education), 
             data = training,
             family = bernoulli(),

                   )
```

```{r}
testing <- data[boundary:total_rows,]

probability <- predict(model, newdata = training, "probs")
column_index <- max.col(probability, tie="random")

testing_predictions <- colnames(probability)[column_index]
testing_ground_truth <- training %>% pull(vote_2020)
accuracy(testing_ground_truth, testing_predictions)

```

```{r}
table(reduced_data$state)
table(reduced_data$race_ethnicity)
table(reduced_data$age)
table(reduced_data$gender)
table(reduced_data$household_income)

table(reduced_data$vote_intention)
table(reduced_data$vote_2020)
```
```{r}
model <- nnet::multinom(vote_2020 ~ state + race_ethnicity + gender + 
                          household_income, data = reduced_data)
model2 <- nnet::multinom(vote_2020 ~ household_income, data = reduced_data)
```
```{r}
summary(model)
```

```{r}
pre
```