### 2020-10-29 multi-model.Rmd

---
title: "ps4_models"
author: "James Bao"
date: "10/27/2020"
output: pdf_document
---

```{r}
library(haven)
library(tidyverse)
library(nnet)
library(Metrics)

```

<!-- ```{r} -->
<!-- raw_data <- read_dta("inputs/data/usa_00001.dta") -->
<!-- # Add the labels -->
<!-- raw_data <- labelled::to_factor(raw_data) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # Just keep some variables that may be of interest (change  -->
<!-- # this depending on your interests) -->
<!-- names(raw_data) -->

<!-- reduced_data <-  -->
<!--   raw_data %>%  -->
<!--   select( -->
<!--          stateicp, -->
<!--          sex, -->
<!--          race,  -->
<!--          educd, -->
<!--          inctot) -->

<!-- head(reduced_data) -->
<!-- reduced_data$race %>% table() -->
<!-- reduced_data$stateicp %>% table() -->
<!-- reduced_data$educd %>% table() -->
<!-- reduced_data$sex %>% table() -->
<!-- reduced_data$inctot %>% table() -->

<!-- strat <- reduced_data -->
<!-- ``` -->

```{r}
raw_data <- read_dta("inputs/data/ns20200625/ns20200625.dta")
# Add the labels
raw_data <- labelled::to_factor(raw_data)
```
```{r}
colnames(raw_data)
```

```{r}
reduced_data <- 
  raw_data %>% 
  select(vote_intention,
         vote_2020,
         state,
         race_ethnicity,
         age,
         gender,
         household_income,
         education
  )
reduced_data <- reduced_data[complete.cases(reduced_data), ]

reduced_data$race_ethnicity <- as.character(reduced_data$race_ethnicity)
reduced_data$race_ethnicity[
    reduced_data$race_ethnicity == 'Asian (Asian Indian)' |
    reduced_data$race_ethnicity == 'Asian (Korean)' |
    reduced_data$race_ethnicity == 'Asian (Filipino)' |
    reduced_data$race_ethnicity == 'Asian (Vietnamese)'|
    grepl("Pacific", reduced_data$race_ethnicity)
  ] <- ('Asian (Other)')
reduced_data$race_ethnicity <- as.factor(reduced_data$race_ethnicity)

total_rows <- nrow(reduced_data)

```
```{r}
shuffle_indices <- sample(total_rows)
data <- reduced_data[shuffle_indices,]
boundary <- as.integer(total_rows*0.9)

training <- data[0:boundary,]
model <- nnet::multinom(vote_2020 ~ state + race_ethnicity + gender + 
                          education, data = training)
```

```{r}
testing <- data[boundary:total_rows,]

probability <- predict(model, newdata = testing, "probs")
column_index <- max.col(probability, tie="random")

testing_predictions <- colnames(probability)[column_index]
testing_ground_truth <- testing %>% pull(vote_2020)
accuracy(testing_ground_truth, testing_predictions)

```

```{r}
table(reduced_data$state)
table(reduced_data$race_ethnicity)
table(reduced_data$age)
table(reduced_data$gender)
table(reduced_data$household_income)

table(reduced_data$vote_intention)
table(reduced_data$vote_2020)
```
```{r}
model <- nnet::multinom(vote_2020 ~ state + race_ethnicity + gender + 
                          household_income, data = reduced_data)
model2 <- nnet::multinom(vote_2020 ~ household_income, data = reduced_data)
```
```{r}
summary(model)
```

```{r}
pre
```