### 2020-11-02 predict.Rmd
more stuff

---
title: "predict"
author: "James Bao"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
cell_counts <- readRDS("inputs/cell_counts.RDS")
model <- readRDS("model/4chains_3000iter_.rds")
```

```{r}
overall_prop <- cell_counts %>%
  mutate(prop = n/sum(n))

overall_prop

```

```{r}
overall_prop$predict <- predict(model, type = "response", newdata = overall_prop)
overall_prop$predict <- overall_prop$predict[,1] 
head(overall_prop$predict)
```

```{r}
overall_prop$vote <- if_else(overall_prop$predict >0.5, 1, 0)
```

```{r}
overall_prop$vote_weight <- overall_prop$vote * overall_prop$prop

```

```{r}
sum(overall_prop$vote_weight)
saveRDS(overall_prop, file='processed_data/predictions.RDS')
```