### 2020-11-02 mrp.Rmd

---
title: "mrp"
author: "James Bao"
date: "10/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(brms)
library(tidybayes)
library(statebins)
library(dplyr)
```

```{r}
#load the necessary model and data frames
cell_counts <- readRDS("inputs/cell_counts.RDS")
training_data <- readRDS("inputs/training_data.RDA")
model <- readRDS("model/4chains_3000iter_.rds")

```

```{r}
#load the summarized results for each variable we used
race_res <- readRDS("processed_data/race_res.RDS")
gender_res <- readRDS("processed_data/gender_res.RDS")
state_res <- readRDS("processed_data/state_res.RDS")
age_res <- readRDS("processed_data/age_res.RDS")
education_rest <- readRDS("processed_data/age_res.RDS")

```

```{r}
age_res %>% 
  ggplot(aes(y = mean, x = forcats::fct_inorder(age_group), color = "MRP estimate")) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) + 
  ylab("proportion keeping name") + 
  xlab("age in 2020") + 
  geom_point(data = training_data %>% 
               group_by(age_group, vote_biden) %>%
               dplyr::summarise(n = n()) %>% 
               group_by(age_group) %>% 
               mutate(prop = n/sum(n)) %>% 
               filter(vote_biden==1), 
             aes(age_group, prop, color = "MNCS raw data")) +
  scale_color_manual(name = "", values = c("MRP estimate" = "black", "MNCS raw data" = "red")) + 
  theme_bw(base_size = 14) 
```

```{r}
race_res %>% 
  ggplot(aes(y = mean, x = forcats::fct_inorder(race_ethnicity), color = "MRP estimate")) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) + 
  ylab("proportion keeping name") + 
  xlab("age in 2020") + 
  geom_point(data = training_data %>% 
               group_by(race_ethnicity, vote_biden) %>%
               dplyr::summarise(n = n()) %>% 
               group_by(race_ethnicity) %>% 
               mutate(prop = n/sum(n)) %>% 
               filter(vote_biden==1), 
             aes(race_ethnicity, prop, color = "MNCS raw data")) +
  scale_color_manual(name = "", values = c("MRP estimate" = "black", "MNCS raw data" = "red")) + 
  theme_bw(base_size = 14) 
```

```{r}

state_res %>% 
  ggplot(aes(y = mean, x = forcats::fct_inorder(state), color = "MRP estimate")) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) + 
  ylab("proportion keeping name") + 
  xlab("age in 2020") + 
  geom_point(data = training_data %>% 
               group_by(state, vote_biden) %>%
               dplyr::summarise(n = n()) %>% 
               group_by(state) %>% 
               mutate(prop = n/sum(n)) %>% 
               filter(vote_biden==1), 
             aes(state, prop, color = "MNCS raw data")) +
  scale_color_manual(name = "", values = c("MRP estimate" = "black", "MNCS raw data" = "red")) + 
  theme_bw() +
  coord_flip()

```

```{r}
#show selected states for more clarity
#include key, swing and heavy trump/heavy biden states for discussion

filtered_state_res <- state_res %>% filter(state == "PA" | state == "VT" | 
                                           state == "MI" | state == "AR" |
                                          state == "WI")
filtered_training_data <- training_data %>% filter(state == "PA" | state == "VT" | 
                                           state == "MI" | state == "AR" |
                                          state == "WI")
filtered_state_res %>% 
  ggplot(aes(y = mean, x = forcats::fct_inorder(state), color = "MRP estimate")) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) + 
  ylab("proportion keeping name") + 
  xlab("age in 2020") + 
  geom_point(data = filtered_training_data %>% 
               group_by(state, vote_biden) %>%
               dplyr::summarise(n = n()) %>% 
               group_by(state) %>% 
               mutate(prop = n/sum(n)) %>% 
               filter(vote_biden==1), 
             aes(state, prop, color = "MNCS raw data")) +
  scale_color_manual(name = "", values = c("MRP estimate" = "black", "MNCS raw data" = "red")) + 
  theme_bw() +
  coord_flip()
```

```{r}
gender_res %>% 
  ggplot(aes(y = mean, x = forcats::fct_inorder(gender), color = "MRP estimate")) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) + 
  ylab("proportion keeping name") + 
  xlab("Gender in 2020") + 
  geom_point(data = training_data %>% 
               group_by(gender, vote_biden) %>%
               dplyr::summarise(n = n()) %>% 
               group_by(gender) %>% 
               mutate(prop = n/sum(n)) %>% 
               filter(vote_biden==1), 
             aes(gender, prop, color = "MNCS raw data")) +
  scale_color_manual(name = "", values = c("MRP estimate" = "black", "MNCS raw data" = "red")) + 
  theme_bw(base_size = 14) 
```

```{r}
education_res %>% 
  ggplot(aes(y = mean, x = forcats::fct_inorder(education), color = "MRP estimate")) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) + 
  ylab("proportion keeping name") + 
  xlab("Gender in 2020") + 
  geom_point(data = training_data %>% 
               group_by(education, vote_biden) %>%
               dplyr::summarise(n = n()) %>% 
               group_by(education) %>% 
               mutate(prop = n/sum(n)) %>% 
               filter(vote_biden==1), 
             aes(education, prop, color = "MNCS raw data")) +
  scale_color_manual(name = "", values = c("MRP estimate" = "black", "MNCS raw data" = "red")) + 
  theme_bw(base_size = 14) 
```